---
title: "NorthRoss_MetaAnalysis"
output: html_document
---

Clean up data

```{r}
library(metafor)
library(orchaRd)
df <- read.csv("ExtractedData_2026-05-15.csv")
# Convert all variances to SD
table(df$Biodiversity_Variance_unit)
df$Control_SD = df$Control_var
df$Treatment_SD = df$Treatment_var
for(i in 1:nrow(df)){
  dfi <- df[i,]
  if(dfi$Biodiversity_Variance_unit == "SE"){

    dfi$Control_SD = dfi$Control_var * sqrt(dfi$Control_n)
    dfi$Treatment_SD = dfi$Treatment_var * sqrt(dfi$Treatment_n)
  }
  df[i,] <- dfi
}

# Note -some of the variances are too high and need to be corrected

# Combine peatland types
df$Peatland_type <- replace(df$Peatland_type, 
                            grepl(" fen", df$Peatland_type),
                            "Fen")
table(df$Peatland_type)

# Combine functional groups
df$FG_Group <- replace(df$FG_Group, 
                      grepl(" moss", df$FG_Group),
                      "Moss")

# Convert some variables to factors
df$Peatland_type <- factor(df$Peatland_type)
df$FG_Group <- factor(df$FG_Group)

# Convert to numeric
df$WTD_Difference <- as.numeric(df$WTD_Difference)
```

# Func Groups as mod

```{r}
# Let's try doing all functional group data then separating them as a variable
fg <- subset(df, df$SR_or_FG=="FG")

# Remove empty
fg <- subset(fg, fg$FG_Group != "")
# Remove "vascular plant" (not really useful for comparison)
fg <- subset(fg, fg$FG_Group != "Vascular plant")

fg$FG_Group <- factor(fg$FG_Group)
# Get effect size
fgES <- escalc(measure = "SMDH", # I think this is best, 
                # since many of the studies use different units and 
                # measurement techniques and different sample sizes
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = fg)
# Try a one-variable with FG
fg1 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~FG_Group, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg1)
# only significant change is in trees at p=0.01

orchard_plot(fg1,
             mod = "FG_Group",
             group = "PaperID",
             xlab = "Standardised mean difference (Treatment - Control) with heteroscedastic population variances")
# I think this looks pretty good?
```

Only one FG changed significantly

```{r}

# Try with WTD difference as another mod
fg2 <-rma.mv(
  yi=yi,
  V=vi,
  mods = ~ FG_Group * WTD_Difference,
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg2)



fg2_plot <- mod_results(fg2, mod = "WTD_Difference", group = "PaperID", by = "FG_Group")

bubble_plot(fg2_plot, group = "PaperID",  mod = "WTD_Difference", xlab = "WTD Difference")
```

WTD on its own has a significant negative effect on overall abundance - productivity in general (regardless of functional group) is lower under water table draw down (not what we would expect), but this is probably biased by differing methods and measurements.

Where is "Herb" in the summary????

# Looking at just herbs with peatland type as mod

```{r}
# Filter to just herbs
# Later I will compare the effect on herbs, trees and mosses
table(df$FG_Group)
herbs <- subset(df,df$FG_Group=="Herb")

# Get effect size
datES <- escalc(measure = "SMDH", # I think this is best, 
                    # since many of the studies use different units and 
                    # measurement techniques and different sample sizes
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = herbs)

# Grand mean
grandmean1 <-rma.mv(
  yi=yi,
  V=vi, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=datES)
summary(grandmean1)
# Try a one-variable with Peatland type
peat1 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~Peatland_type, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=datES)

orchard_plot(peat1,
             mod = "Peatland_type",
             group = "PaperID",
             xlab = "Standardised mean difference (with heteroscedastic population variances)")
```

The grand mean of the effect size was -0.2296, but only had a p-value of 0.1946, indicating that there was no significant effect of WLD on herb production.
