---
title: "NorthRoss_MetaAnalysis"
output:
  pdf_document: default
  html_document: default
---

Clean up data

```{r}
library(metafor)
library(orchaRd)
library(tidyr)

df <- read.csv("ExtractedData_2026-05-31.csv")
# Convert all variances to SD
table(df$Biodiversity_Variance_unit)
df$Control_SD = df$Control_var
df$Treatment_SD = df$Treatment_var
for(i in 1:nrow(df)){
  dfi <- df[i,]
  if(dfi$Biodiversity_Variance_unit == "SE"){

    dfi$Control_SD = dfi$Control_var * sqrt(dfi$Control_n)
    dfi$Treatment_SD = dfi$Treatment_var * sqrt(dfi$Treatment_n)
  }
  df[i,] <- dfi
}

# remove record if NAs in any of these columns:
fieldsList = c("Control_mean", "Control_n", "Control_var",
               "Treatment_mean", "Treatment_n", "Treatment_var")
for (i in 1:6){
  df <- drop_na(df, fieldsList[i])
}
  

# Combine peatland types
df$Peatland_type <- replace(df$Peatland_type, 
                            grepl(" fen", df$Peatland_type),
                            "Fen")
table(df$Peatland_type)

# Combine functional groups
df$FG_Group <- replace(df$FG_Group, 
                      grepl(" moss", df$FG_Group),
                      "Moss")

# Convert some variables to factors
df$Peatland_type <- factor(df$Peatland_type)
df$FG_Group <- factor(df$FG_Group)

# Convert to numeric
df$WTD_Difference <- as.numeric(df$WTD_Difference)
```

# Func Groups as mod

```{r}
# Let's try doing all functional group data then separating them as a variable
fg <- subset(df, df$SR_or_FG=="FG")

# Remove empty
fg <- subset(fg, fg$FG_Group != "")
# Remove "vascular plant" (not really useful for comparison)
fg <- subset(fg, fg$FG_Group != "Vascular plant")

fg$FG_Group <- factor(fg$FG_Group)
# Get effect size
fgES <- escalc(measure = "ROM", # I think this is best due to the variety of units used
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = fg)
# Try a one-variable with FG
fg1 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~FG_Group -1, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg1)
# only significant change is in trees at p=0.01

I2 <- orchaRd::i2_ml(fg1, method="matrix")

orchard_plot(fg1,
             mod = "FG_Group",
             group = "PaperID",
             xlab = "Log response ratio (ROM)")
```

Highly significant increase in trees, decrease in moss at p=0.1. Non-significant decrease in herbs (unexpected) and increase in shrubs (expected).

Try using CaseID as a random effect and group in plot:

```{r}
# Try a one-variable with FG
fg3 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~FG_Group -1, 
  random= ~1|PaperID/Case_ID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg3)
# only significant change is in trees at p=0.01

orchard_plot(fg3,
             mod = "FG_Group",
             group = "Case_ID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")
```

This makes things more significant, but it might be due to overfitting since we don't have very many data points for separate cases within each paper.

The confidence intervals are on both sides of zero in all groups, meaning that the data is highly heterogeneous. This is also indicated by the Test for Residual Heterogeneity ($Q_E$ test) in the summary which has a significant result. Now would be a good time to check if our results are being skewed by outliers, but I don't have time to dig into this. From the orchard plot, it appears the results are fairly uniform.

```{r}

# Try with WTD difference as another mod
fg2 <-rma.mv(
  yi=yi,
  V=vi,
  mods = ~ FG_Group * WTD_Difference -1,
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg2)



fg2_plot <- mod_results(fg2, mod = "WTD_Difference", group = "PaperID", by = "FG_Group")

bubble_plot(fg2_plot, group = "PaperID",  mod = "WTD_Difference", xlab = "WTD_Difference")

```

Herb production was significantly negatively affected by WTD difference at the p=0.01 level - studies with small changes in WTD resulted in slight decreases in herb abundance,while high WTD showed increases. This makes sense - studies that didn't really change WTD much didn't have a strong effect on shrubs, but large decreases in WTD had increased shrub production. (note that "WTD_Difference" above is actually FG_GroupHerb:WTD_Difference)

Mosses also had a significant increase at the p=0.01 level, but it's a very small increase (0.0765 units/cm), doesn't really tell us much. Herbs were also significant

# Species Richness

We didn't get a lot of data on species richness, but let's take a look at it:

```{r}
sr <- subset(df, df$SR_or_FG=="SR")
# Remove NA
sr <- subset(sr, !is.na(sr$Control_mean))

# Remove the Koster 2023 records with zero variance (shouldn't have been used)
sr <- subset(sr, !PaperID == 12)
# And the Andrew ones, I don't think these are appropriate
sr <- subset(sr, !PaperID == 67)

print(length(sr[,1]))
print(length(unique(sr$PaperID)))
# only 5 records from 3 papers...

# factorize peatland type
sr$Peatland_type <- factor(sr$Peatland_type)

# Get effect size
srES <- escalc(measure = "SMDH", # Since this time they have consistent units
               # standardized mean difference with heteroscedastic population variances
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = sr)
# grand mean
sr1 <-rma.mv(
  yi=yi,
  V=vi, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr1)
# mean slight increase in species richness
# but no significant difference
orchard_plot(sr1,
             group = "PaperID",
             xlab = "Species richness standardised mean difference")

```

Not significant, but overall the grand mean is negative, meaning that the

```{r}
# Try a one-variable with peatland type
# Although we don't really have enough data for this
sr2 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~Peatland_type -1, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr2)
# No significant effects
# bogs tended to decrease in SR with drawdown while other peatlands increased

orchard_plot(sr2,
             mod = "Peatland_type",
             group = "PaperID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")
```

```{r}
# Try continuous variable WTD difference
# note that one study didn't report WTD
# Try with WTD difference as another mod
sr3 <-rma.mv(
  yi=yi,
  V=vi,
  mods = ~ WTD_Difference,
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr3)
# no significant effect
# smaller differences in WTD had different 

sr3_plot <- mod_results(sr3, mod = "WTD_Difference", group = "PaperID")
bubble_plot(sr3_plot, group = "PaperID",  mod = "WTD_Difference", xlab = "WTD_Difference")
```

# TO DO:

-   Find outliers (<https://doing-meta.guide/heterogeneity.html?q=case#outliers>)

-   
