---
title: "NorthRoss_MetaAnalysis"
output:
  pdf_document: default
  html_document: default
---

Clean up data

```{r}
library(metafor)
library(orchaRd)
library(tidyr)

df <- read.csv("ExtractedData_2026-05-31.csv")
# Convert all variances to SD
table(df$Biodiversity_Variance_unit)
df$Control_SD = df$Control_var
df$Treatment_SD = df$Treatment_var
for(i in 1:nrow(df)){
  dfi <- df[i,]
  if(dfi$Biodiversity_Variance_unit == "SE"){

    dfi$Control_SD = dfi$Control_var * sqrt(dfi$Control_n)
    dfi$Treatment_SD = dfi$Treatment_var * sqrt(dfi$Treatment_n)
  }
  df[i,] <- dfi
}

# remove record if NAs in any of these columns:
fieldsList = c("Control_mean", "Control_n", "Control_var",
               "Treatment_mean", "Treatment_n", "Treatment_var")
for (i in 1:6){
  df <- drop_na(df, fieldsList[i])
}
  

# Combine peatland types
df$Peatland_type <- replace(df$Peatland_type, 
                            grepl(" fen", df$Peatland_type),
                            "Fen")
table(df$Peatland_type)

# Combine functional groups
df$FG_Group <- replace(df$FG_Group, 
                      grepl(" moss", df$FG_Group),
                      "Moss")

# Convert some variables to factors
df$Peatland_type <- factor(df$Peatland_type)
df$FG_Group <- factor(df$FG_Group)

# Convert to numeric
df$WTD_Difference <- as.numeric(df$WTD_Difference)
```

# Func Groups as mod

```{r}
# Let's try doing all functional group data then separating them as a variable
fg <- subset(df, df$SR_or_FG=="FG")

# Remove empty
fg <- subset(fg, fg$FG_Group != "")
# Remove "vascular plant" (not really useful for comparison)
fg <- subset(fg, fg$FG_Group != "Vascular plant")

fg$FG_Group <- factor(fg$FG_Group)
# Get effect size
fgES <- escalc(measure = "SMDH", # I think this is best, 
                # since many of the studies use different units and 
                # measurement techniques and different sample sizes
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = fg)
# Try a one-variable with FG
fg1 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~FG_Group -1, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg1)
# only significant change is in trees at p=0.01

orchard_plot(fg1,
             mod = "FG_Group",
             group = "PaperID",
             xlab = "Effect size (standardised mean difference, control - treatment)")
```

Highly significant increase in trees, decrease in moss at p=0.1. Non-significant decrease in herbs (unexpected) and increase in shrubs (expected).

Try using CaseID as a random effect and group in plot:

```{r}
# Try a one-variable with FG
fg3 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~FG_Group -1, 
  random= ~1|PaperID/Case_ID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg3)
# only significant change is in trees at p=0.01

orchard_plot(fg3,
             mod = "FG_Group",
             group = "Case_ID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")
```

This makes things more significant, but it might be due to overfitting since we don't have very many data points for separate cases within each paper.

The confidence intervals are on both sides of zero in all groups, meaning that the data is highly heterogeneous. This is also indicated by the Test for Residual Heterogeneity ($Q_E$ test) in the summary which has a significant result. Let's see if this is being affected by outliers:

```{r}
# add dmetar library to get find.outliers tools
library(devtools)
devtools::install_github("MathiasHarrer/dmetar")
library(dmetar)
```

```{r}

# Try with WTD difference as another mod
fg2 <-rma.mv(
  yi=yi,
  V=vi,
  mods = ~ FG_Group * WTD_Difference -1,
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg2)



fg2_plot <- mod_results(fg2, mod = "WTD_Difference", group = "PaperID", by = "FG_Group")

bubble_plot(fg2_plot, group = "PaperID",  mod = "WTD_Difference", xlab = "WTD_Difference")

fg2_plot <- mod_results(fg2, mod = "WTD_Difference", group = "PaperID", by = "FG_Group")
bubble_plot(fg2_plot, group = "PaperID",  mod = "Time_frame", xlab = "Time_frame")
```

Herb production was significantly negatively affected by WTD difference at the p=0.01 level - studies with small changes in WTD resulted in slight decreases in herb abundance,while high WTD showed increases. This makes sense - studies that didn't really change WTD much didn't have a strong effect on shrubs, but large decreases in WTD had increased shrub production. (note that "WTD_Difference" above is actually FG_GroupHerb:WTD_Difference)

Mosses also had a significant increase at the p=0.01 level, but it's a very small increase (0.0765 units/cm), doesn't really tell us much. Herbs were also significant

# Looking at just herbs with peatland type as mod

```{r}
# Filter to just herbs
# Later I will compare the effect on herbs, trees and mosses
table(df$FG_Group)
herbs <- subset(df,df$FG_Group=="Herb")

# Get effect size
datES <- escalc(measure = "SMDH", # I think this is best, 
                    # since many of the studies use different units and 
                    # measurement techniques and different sample sizes
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = herbs)

# Grand mean
grandmean1 <-rma.mv(
  yi=yi,
  V=vi, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=datES)
summary(grandmean1)
# Try a one-variable with Peatland type
peat1 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~Peatland_type, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=datES)

orchard_plot(peat1,
             mod = "Peatland_type",
             group = "PaperID",
             xlab = "Standardised mean difference (with heteroscedastic population variances)")
```

The grand mean of the effect size was -0.2296, but only had a p-value of 0.1946, indicating that there was no significant effect of WLD on herb production.

# Species Richness

We didn't get a lot of data on species richness, but let's take a look at it:

```{r}
sr <- subset(df, df$SR_or_FG=="SR")
# Remove NA
sr <- subset(sr, !is.na(sr$Control_mean))

print(length(sr[,1]))
# only 9 records...

# factorize peatland type
sr$Peatland_type <- factor(sr$Peatland_type)

# Get effect size
srES <- escalc(measure = "ROM", # Since we are looking at positive numbers similar in scale to each other
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = sr)
# grand mean
sr1 <-rma.mv(
  yi=yi,
  V=vi, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr1)
# no significant difference
orchard_plot(sr1,
             group = "PaperID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")

```

Not significant, but overall the grand mean is negative, meaning that the

```{r}
# Try a one-variable with peatland type
# Although we don't really have enough data for this
sr2 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~Peatland_type -1, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr2)
# No significant effects
# bogs tended to decrease in SR with drawdown while other peatlands increased

orchard_plot(sr2,
             mod = "Peatland_type",
             group = "PaperID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")
```

```{r}
# Try continuous variable WTD difference
# note that one study didn't report WTD
# Try with WTD difference as another mod
sr3 <-rma.mv(
  yi=yi,
  V=vi,
  mods = ~ WTD_Difference,
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr3)
# no significant effect
# smaller differences in WTD had different 

sr3_plot <- mod_results(sr3, mod = "WTD_Difference", group = "PaperID")
bubble_plot(sr3_plot, group = "PaperID",  mod = "WTD_Difference", xlab = "WTD_Difference")
```

# TO DO:

-   Find outliers (<https://doing-meta.guide/heterogeneity.html?q=case#outliers>)

-   
