---
title: "NorthRoss_MetaAnalysis"
output: html_document
---

Clean up data

```{r}
library(metafor)
library(orchaRd)
df <- read.csv("ExtractedData_2026-05-15.csv")
# Convert all variances to SD
table(df$Biodiversity_Variance_unit)
df$Control_SD = df$Control_var
df$Treatment_SD = df$Treatment_var
for(i in 1:nrow(df)){
  dfi <- df[i,]
  if(dfi$Biodiversity_Variance_unit == "SE"){

    dfi$Control_SD = dfi$Control_var * sqrt(dfi$Control_n)
    dfi$Treatment_SD = dfi$Treatment_var * sqrt(dfi$Treatment_n)
  }
  df[i,] <- dfi
}

# Note -some of the variances are too high and need to be corrected

# Combine peatland types
df$Peatland_type <- replace(df$Peatland_type, 
                            grepl(" fen", df$Peatland_type),
                            "Fen")
table(df$Peatland_type)

# Combine functional groups
df$FG_Group <- replace(df$FG_Group, 
                      grepl(" moss", df$FG_Group),
                      "Moss")

# Convert some variables to factors
df$Peatland_type <- factor(df$Peatland_type)
df$FG_Group <- factor(df$FG_Group)

# Convert to numeric
df$WTD_Difference <- as.numeric(df$WTD_Difference)
```

# Func Groups as mod

```{r}
# Let's try doing all functional group data then separating them as a variable
fg <- subset(df, df$SR_or_FG=="FG")

# Remove empty
fg <- subset(fg, fg$FG_Group != "")
# Remove "vascular plant" (not really useful for comparison)
fg <- subset(fg, fg$FG_Group != "Vascular plant")

fg$FG_Group <- factor(fg$FG_Group)
# Get effect size
fgES <- escalc(measure = "SMDH", # I think this is best, 
                # since many of the studies use different units and 
                # measurement techniques and different sample sizes
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = fg)
# Try a one-variable with FG
fg1 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~FG_Group -1, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg1)
# only significant change is in trees at p=0.01

orchard_plot(fg1,
             mod = "FG_Group",
             group = "PaperID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")
```

No significant changes to

```{r}

# Try with WTD difference as another mod
fg2 <-rma.mv(
  yi=yi,
  V=vi,
  mods = ~ FG_Group * WTD_Difference -1,
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=fgES)
summary(fg2)



fg2_plot <- mod_results(fg2, mod = "WTD_Difference", group = "PaperID", by = "FG_Group")

bubble_plot(fg2_plot, group = "PaperID",  mod = "WTD_Difference", xlab = "WTD_Difference")
```

Herb production was significantly affected by WTD difference at the p=0.05 level - studies with small changes in WTD resulted in decreases in herb abundance,while high WTD showed increases.

# Looking at just herbs with peatland type as mod

```{r}
# Filter to just herbs
# Later I will compare the effect on herbs, trees and mosses
table(df$FG_Group)
herbs <- subset(df,df$FG_Group=="Herb")

# Get effect size
datES <- escalc(measure = "SMDH", # I think this is best, 
                    # since many of the studies use different units and 
                    # measurement techniques and different sample sizes
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = herbs)

# Grand mean
grandmean1 <-rma.mv(
  yi=yi,
  V=vi, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=datES)
summary(grandmean1)
# Try a one-variable with Peatland type
peat1 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~Peatland_type, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=datES)

orchard_plot(peat1,
             mod = "Peatland_type",
             group = "PaperID",
             xlab = "Standardised mean difference (with heteroscedastic population variances)")
```

The grand mean of the effect size was -0.2296, but only had a p-value of 0.1946, indicating that there was no significant effect of WLD on herb production.

# Species Richness

We didn't get a lot of data on species richness, but let's take a look at it:

```{r}
sr <- subset(df, df$SR_or_FG=="SR")
# Remove NA
sr <- subset(sr, !is.na(sr$Control_mean))
# factorize peatland type
sr$Peatland_type <- factor(sr$Peatland_type)

# Get effect size
srES <- escalc(measure = "ROM", # Since we are looking at positive numbers similar in scale to each other
                m2i = Control_mean, 
                sd2i = Control_SD,
                n2i = Control_n,
                m1i = Treatment_mean,
                sd1i = Treatment_SD,
                n1i = Treatment_n,
                data = sr)
# grand mean
sr1 <-rma.mv(
  yi=yi,
  V=vi, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr1)
# no significant difference
orchard_plot(sr1,
             group = "PaperID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")

```

```{r}
# Try a one-variable with peatland type
# Although we don't really have enough data for this
sr2 <-rma.mv(
  yi=yi,
  V=vi, 
  mods=~Peatland_type -1, 
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr2)
# No significant effects
# bogs tended to decrease in SR with drawdown while other peatlands increased

orchard_plot(sr2,
             mod = "Peatland_type",
             group = "PaperID",
             xlab = "Standardised mean difference (Control - Treatmet) with heteroscedastic population variances")
```

```{r}
# Try continuous variable WTD difference
# note that one study didn't report WTD
# Try with WTD difference as another mod
sr3 <-rma.mv(
  yi=yi,
  V=vi,
  mods = ~ WTD_Difference,
  random= ~1|PaperID/ID,
  method="REML",
  digits=4,
  data=srES)
summary(sr3)
# no significant effect
# smaller differences in WTD had different 

sr3_plot <- mod_results(sr3, mod = "WTD_Difference", group = "PaperID")
bubble_plot(sr3_plot, group = "PaperID",  mod = "WTD_Difference", xlab = "WTD_Difference")
```
